---
title: "Week"
author: "Felix Lanver"
date: "2023-01-13"
output: html_document
---

# Set Up and Data

```{r setup}

#packages
pacman::p_load(tidyverse, ggfx, patchwork, showtext, mapproj, broom)


# Get the Data
feederwatch <- tidytuesdayR::tt_load(2023, week = 2)
# get public & countsite data
public <- feederwatch$PFW_2021_public
countsite <- feederwatch$PFW_count_site_data_public_2021

# remove Canada, Alaska & Hawaii the simplify map
public_us <- public |>
  filter(grepl("US-", subnational1_code),
         subnational1_code != "US-AK",
         subnational1_code != "US-HI")

# Import fonts (not sure why, maybe change this later to be a bit more idnividual)

font_add_google(name = "Roboto Slab", family = "roboto slab")
showtext_auto()

# colors
neg <- "#EB5353"
pos <- "#36AE7C"
bg <- "#009999"


```


```{r data}



```



```{r data wrangling}
# prepare data for state division

states.names <- tibble(
  code = state.abb,
  name = state.name
)
```



```{r}
# filter data to only include red-tailed hawk (by state)
rethaw <- public_us %>% 
  filter(species_code == "rethaw", valid == 1) %>% 
  separate(subnational1_code, into = c("country", "state"), sep = "-") %>% 
  filter(!country %in% c("XX", "CA")) %>% 
  left_join(states.names, by = c("state" = "code")) %>% 
  mutate(name = case_when(state == "DC" ~ "District of Columbia",
                          TRUE ~ name)) %>% 
  mutate(name = tolower(name)) %>% 
  select(us_state = name, Year, Month, Day, how_many) %>% 
  arrange(Year, Month, Day)

# 
count_per_state <- rethaw %>% 
  group_by(us_state) %>% 
  summarise(total = sum(how_many))

map_data <- map_data("state") %>% 
  left_join(count_per_state, by = c("region" = "us_state"))

# df for sightings per month
plot1_data <- rethaw %>% 
  group_by(Year, Month) %>% 
  summarise(total = sum(how_many)) %>% 
  rowid_to_column()

# df for sightings per day
plot2_data <- rethaw %>% 
  mutate(date = lubridate::as_date(paste(Year, Month, Day, sep = "-"))) %>% 
  mutate(weekday = lubridate::wday(date, abbr = FALSE, label = TRUE, week_start = 7, locale = "en")) %>% 
  group_by(weekday) %>% 
  summarise(total = sum(how_many)) %>% 
  mutate(weekday = fct_rev(fct_inorder(weekday)))

rm(count_per_state, feederwatch, states.names)
```



```{r map}

map <- ggplot(data = map_data,
         aes(x = long, y = lat, group = group, fill = total)) +
  geom_polygon(col = "grey70",
               show.legend = FALSE) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  scale_fill_gradient2(low = "#ffcccc",
                       mid = "#ff9999",
                       high = "#ff6666",
                       midpoint = max(map_data$total, na.rm = TRUE) / 2,
                       na.value = "#cccccc") +
  labs(title = "Frequency of the Species",
       subtitle = "Darker colours indicate more birds") +
  theme_void() +
  theme(panel.background = element_rect(fill = "#009999", colour = "#009999"),
        plot.background = element_rect(fill = "#009999", colour = "#009999"),
        plot.title = element_text(family = "roboto slab", colour = "white",
                                  size = 50, hjust = 0.5,
                                  margin = margin(t = 35)),
        plot.subtitle = element_text(family = "roboto slab", colour = "white",
                                     size = 30, hjust = 0.5))

map
```

```{r plot 1 habitats}
habitat <- countsite %>%
  select(loc_id, "hab_dcid_woods":"hab_marsh") %>%
  drop_na() %>%
  distinct()

model_data <- sightings_data %>%
  clean_names() %>% 
  #filter(valid == 1) %>%
  left_join(habitat, by = "loc_id") %>%
  select(obs_id, eursta_sightings, loc_id, "hab_dcid_woods":"hab_marsh") %>%
  group_by(obs_id) %>%
  mutate(across("hab_dcid_woods":"hab_marsh", ~ max(.x))) %>% 
  ungroup() %>%
  distinct() %>% 
  select(eursta_sightings, hab_dcid_woods:hab_marsh)

model_summary <- lm(eursta_sightings ~ ., data = model_data) %>%
  tidy() %>%
  slice(-1) %>%
  mutate(abs_estimate = abs(estimate),
         is_positive = (estimate > 0),
         term = str_sub(term, 5),
         term = str_replace(term, "_", " "))
  
# plot ---- 

showtext_auto()
font_add_google("Cormorant")
main_font <- "Cormorant"

# define x0
x0 <- max(model_summary$abs_estimate)

title <- "Which habitats attract the most birds?"
subtitle <- "Mean effect of habitat type on number of observed birds in descending order"


p <- model_summary %>%
  ggplot() + 
  ggtitle(title,
         subtitle) +
  geom_linerange(aes(y = reorder(term, estimate), 
                     xmin = 0, xmax = estimate,
                     colour = !is_positive),
                     linewidth = 40) +
  scale_color_manual(values = c(pos, neg)) +
  geom_text(
    mapping = aes(x = 0, y = term, label = term),
    hjust = 0.5,
    nudge_y = 0.4,
    size = 4,
    family = main_font) +
  annotate("text", x = 1.25, y = 7,
           label = "#TidyTuesday 10/01/2023 | Data: feederwatch.org | @_jcken",
           family = main_font, angle = 90, color = "grey60") +
  xlab("Number of birds we expect to see above overall mean") +
  ylab("") +
  xlim(-x0, x0) +
  scale_y_discrete(breaks = NULL) +
  theme_minimal() +
  theme(legend.position = "n",
        plot.title = element_text(hjust = 0.5,
                                  family = main_font,
                                  size = 20),
        plot.subtitle = element_text(hjust = 1,
                                     family = main_font,
                                     size = 14),
        text = element_text(family = main_font),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = bg),
        panel.background = element_rect(fill = bg, colour = bg),
        axis.text.x = element_text(size = 16)) 

p
```


# Plot 1 ----

plot1 <- 
  ggplot() +
  geom_rect(aes(xmin = 1, xmax = 2, ymin = -4500, ymax = -2500),
            col = "#ccf9ff", fill = "#251a5c", linewidth = 0.5) +
  geom_text(aes(x = 1.5, y = -3500, label = "2020"),
            col = "#ccf9ff", family = "roboto slab", size = 9) +
  geom_rect(aes(xmin = 2, xmax = 6, ymin = -4500, ymax = -2500),
            col = "#ccf9ff", fill = "#251a5c", linewidth  = 0.5) +
  geom_text(aes(x = 4, y = -3500, label = "2021"),
            col = "#ccf9ff", family = "roboto slab", size = 10) +
  geom_text(aes(x = 1, y = -1500, label = "Nov"),
            col = "#ccf9ff", family = "roboto slab", size = 10) +
  geom_text(aes(x = 2, y = -1500, label = "Dec"),
            col = "#ccf9ff", family = "roboto slab", size = 10) +
  geom_text(aes(x = 3, y = -1500, label = "Jan"),
            col = "#ccf9ff", family = "roboto slab", size = 10) +
  geom_text(aes(x = 4, y = -1500, label = "Feb"),
            col = "#ccf9ff", family = "roboto slab", size = 10) +
  geom_text(aes(x = 5, y = -1500, label = "Mar"),
            col = "#ccf9ff", family = "roboto slab", size = 10) +
  geom_text(aes(x = 6, y = -1500, label = "Apr"),
            col = "#ccf9ff", family = "roboto slab", size = 10) +
  geom_segment(data = plot1_data,
               aes(x = rowid, xend = rowid, y = 500, yend = total),
               linetype = "dotted", col = "white") +
  with_outer_glow(
    geom_line(data = plot1_data,
              aes(x = rowid, y = total),
              col = "#ccf9ff", linewidth = 1.5),
    colour = "#55d0ff",
    sigma = 10) +
  with_outer_glow(
    geom_point(data = plot1_data,
               aes(x = rowid, y = total),
               shape = 21, col = "#ccf9ff", fill = "#251a5c", size = 10, stroke = 1),
    colour = "#55d0ff",
    sigma = 10) +
  geom_text(data = plot1_data,
            aes(x = rowid, y = total, label = total),
            col = "#ccf9ff", size = 10) +
  xlim(c(0, 7)) +
  ylim(c(-4750, 9000)) +
  labs(title = "Number of sightings per month") +
  theme_void() +
  theme(panel.background = element_rect(fill = "#251a5c", colour = "#251a5c"),
        plot.background = element_rect(fill = "#251a5c", colour = "#251a5c"),
        plot.margin = margin(r = 25, b = 10, l = 10),
        plot.title = element_text(family = "roboto slab", colour = "white",
                                  size = 50, hjust = 0.5,
                                  margin = margin(t = 5, b = -20)))

# Plot 2 ----

plot2 <- 
  ggplot() +
  with_outer_glow(
    geom_segment(data = plot2_data,
                 aes(y = weekday, yend = weekday, x = 0, xend = total),
                 col = "#ccf9ff", linewidth = 1),
    colour = "#55d0ff", 
    sigma = 10) +
  geom_text(data = plot2_data,
            aes(x = total + 1250, y = weekday, label = total),
            colour = "#ccf9ff", size = 15) +
  xlim(c(0, 15000)) +
  labs(title = "Number of sightings per weekday",
       x = "", y = "") +
  theme(panel.background = element_rect(fill = "#251a5c", colour = "#251a5c"),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "#251a5c", colour = "#251a5c"),
        plot.margin = margin(r = 25, b = 10, l = 10),
        plot.title = element_text(family = "roboto slab", colour = "white",
                                  size = 50, hjust = 0.5,
                                  margin = margin(t = 15, b = 10)),
        axis.text.x = element_blank(),
        axis.text.y = element_text(family = "roboto slab", colour = "#ccf9ff",
                                   size = 35),
        axis.ticks = element_blank())

# Create patchwork ----

p <- map + (plot1 / plot2) +
  plot_layout(widths = c(2, 1.5)) +
  plot_annotation(
    title = "Sightings of dark-eyed juncos in the U.S.",
    subtitle = "between November 2020 & April 2021",
    caption = "J.Kitt | Data : Project FeederWatch",
    theme = theme(plot.background = element_rect(fill = "#251a5c", colour = "#251a5c"),
                  plot.title = element_text(family = "roboto slab", colour = "white", size = 100,
                                            margin = margin(t = 10), hjust = 0.5),
                  plot.subtitle = element_text(family = "roboto slab", colour = "white", size = 50,
                                               margin = margin(t = 5, b = 10), hjust = 0.5),
                  plot.caption = element_text(family = "roboto slab", colour = "white", size = 30,
                                              margin = margin(t = 5, b = 10), hjust = 0.5)))


# Save plot ----

ggsave("figs/2023_02_feederwatch.png", p, dpi = 320, width = 12, height = 6)